---
title: "Summarising en-net online forum statistics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Summarising en-net online forum statistics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(ennet)
library(dplyr)
```

The `ennet` package also includes analytic functions that summarises the text data available from the [**en-net** online forum](https://www.en-net.org). Currently, there are four sets of analytic functions available from `ennet`:

* `count_topics_time` functions - set of functions to count the number of topics or questions by day, wee, month or year;

* `count_topics_theme` functions - set of functions to count the number of topics or questions by theme;

* `count_topics_author` functions - set of functions to count the number of topics or questions by author; and,

* `arrange` functions - set of functions to arrange topics by number of views or by number of replies.

## Counting number of topics/questions

Summarising the number of topics or questions raised within the [**en-net** online forum](https://www.en-net.org) is basic and useful analytics that can proxy relative importance of a thematic area within the forum. This is facilitated by three sets of functions - 1) `count_topics_time`; 2) `count_topics_theme`; and, 3) `count_topics_author`.

### Counting number of topics/questions by time

The `count_topics_time` set of functions consist of four functions that count topics by *day*, by *week*, by *month*, or by *year*. To count the number of topics/questions per day, the `count_topics_day` function is used as follows:

```{r count_day1, echo = TRUE, eval = FALSE}
count_topics_day()
```

which results in:

```{r count_day1a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  count_topics_day()
```

By default, the `count_topics_day` function will provide an output that is not sorted by number of topics/questions per day. To sort by number of topics/questions per day, specify `.sort` as `TRUE`:

```{r count_day2, echo = TRUE, eval = FALSE}
count_topics_day(.sort = TRUE)
```

which gives the following output:

```{r count_day2a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  count_topics_day(.sort = TRUE)
```

Counting topics/questions by week, month, or year require just using the function named after the time required. For weekly counts, use `count_topics_week`, for monthly use `count_topics_month`, and for yearly use `count_topics_year`.

```{r count_time, echo = TRUE, eval = FALSE}
## Count topics/questions by week
count_topics_week()

## Count topics/questions by month
count_topics_month()

## Count topics/questions by year
count_topics_year()
```

### Counting number of topics/questions by theme

The `count_topics_theme` set of functions consists of two functions that count topics/questions *by theme*, and count topics/questions *by theme and by time*.

Counting of topics by theme is specified as follows:

```{r count_theme1, echo = TRUE, eval = FALSE}
count_topics_theme()
```

which results in:

```{r count_theme1a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  count_topics_theme()
```

Results provided by `count_topics_theme` is not sorted by number of topics/questions. To sort, specify `.sort` as `TRUE`:

```{r count_theme2, echo = TRUE, eval = FALSE}
count_topics_theme(.sort = TRUE)
```

which results in:

```{r count_theme2a, echo = FALSE, eval = TRUE}
count_topics_theme(.sort = TRUE)
```

Counting of topics/questions by theme can also be done by time using the `count_topics_theme_time`. For example, counting topics/questions by theme by day is done as follows:

```{r count_theme_day1, echo = TRUE, eval = FALSE}
count_topics_theme_time(by_time = "day")
```

which gives the following output:

```{r count_theme_day1a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  count_topics_theme_time(by_time = "day")
```

Unlike the previous count functions, the output of `count_topics_theme_time` is already sorted by default.

To count topics/questions by theme by week, month, or year just requires changing the `by_time` argument to the desired time interval as shown below:

```{r count_theme_day2, echo = TRUE, eval = FALSE}
## Count topics/questions by theme by week
count_topics_theme_time(by_time = "week")

## Count topics/questions by theme by month
count_topics_theme_time(by_time = "month")

## Count topics/questions by theme by year
count_topics_theme_time(by_time = "year")
```

### Counting number of topics/questions by author

The `count_topics_author` set of functions consists of two functions that count topics/questions *by author*, and count topics/questions *by author and by time*.

Counting of topics by author is specified as follows:

```{r count_author1, echo = TRUE, eval = FALSE}
count_topics_author()
```

which results in:

```{r count_author1a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  count_topics_author()
```

Results provided by `count_topics_author` is not sorted by number of topics/questions. To sort, specify `.sort` as `TRUE`:

```{r count_author2, echo = TRUE, eval = FALSE}
count_topics_author(.sort = TRUE)
```

which results in:

```{r count_author2a, echo = FALSE, eval = TRUE}
ennet_topics %>% count_topics_author(.sort = TRUE)
```

Counting of topics/questions by author can also be done by time using the `count_topics_author_time`. For example, counting topics/questions by author by day is done as follows:

```{r count_author_day1, echo = TRUE, eval = FALSE}
count_topics_author_time(by_time = "day")
```

which gives the following output:

```{r count_author_day1a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  count_topics_author_time(by_time = "day")
```

Unlike the previous count functions, the output of `count_topics_author_time` is already sorted by default.

To count topics/questions by author by week, month, or year just requires changing the `by_time` argument to the desired time interval as shown below:

```{r count_author_day2, echo = TRUE, eval = FALSE}
## Count topics/questions by author by week
count_topics_author_time(by_time = "week")

## Count topics/questions by author by month
count_topics_author_time(by_time = "month")

## Count topics/questions by author by year
count_topics_author_time(by_time = "year")
```

### Deprecated `count` functions

In the current release of the `ennet` package, the following functions have now undergone deprecation:

* `count_topics`

* `count_authors`

These functions have now been superseded by the more performant `count_topics` functions above. We recommend that if you have been using these deprecated functions previously to update your work to use the new functions described above. These deprecated function will be made defunct on the next release of `ennet` package.

## Arranging topics by number of views

Summarising the number of topics or questions raised within the [**en-net** online forum](https://www.en-net.org) by arranging them based on number of views can proxy level of interest to a specific topic by those participating in the forum. This is facilitated using the `arrange_views` function. Ranking of topics by number of views is done per thematic area and by a specific time period. Ranking of topics by number of views by thematic area and by month and year is performed by default:

```{r usage12, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_views()
```

which results in:

```{r usage12a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_views()
```

Arranging topics by number of views by thematic area and by year is performed as follows:

```{r usage13, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_views(by_date = "year")
```

which results in:

```{r usage13a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_views(by_date = "year")
```

Arranging topics by number of views by thematic area overall across the years is performed as follows:

```{r usage14, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_views(by_date = "all")
```

which results in:

```{r usage14a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_views(by_date = "all")
```

By default, the output of `arrange_views` is grouped by thematic area. This default behaviour can be changed by setting the `by_theme` argument to `FALSE`. For example, to arrange the topics by number of views by month and year across all themes:

```{r usage15, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_views(by_theme = FALSE)
```

which results in:

```{r usage15a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_views(by_theme = FALSE)
```

## Arranging topics by number of replies

Summarising the number of topics or questions raised within the [**en-net** online forum](https://www.en-net.org) by arranging them based on number of replies can proxy level of interest to a specific topic by those participating in the forum specifically those who provide responses and feedback to responses within the discussion. This is facilitated using the `arrange_replies` function. Ranking of topics by number of replies is done per thematic area and by a specific time period. Ranking of topics by number of replies by thematic area and by month and year is performed by default:

```{r usage16, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_replies()
```

which results in:

```{r usage17a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_replies()
```

Arranging topics by number of replies by thematic area and by year is performed as follows:

```{r usage18, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_replies(by_date = "year")
```

which results in:

```{r usage18a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_replies(by_date = "year")
```

Arranging topics by number of replies by thematic area overall across the years is performed as follows:

```{r usage19, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_replies(by_date = "all")
```

which results in:

```{r usage19a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_replies(by_date = "all")
```

By default, the output of `arrange_replies` is grouped by thematic area. This default behaviour can be changed by setting the `by_theme` argument to `FALSE`. For example, to arrange the topics by number of replies by month and year across all themes:

```{r usage20, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_replies(by_theme = FALSE)
```

which results in:

```{r usage20a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_replies(by_theme = FALSE)
```
