---
title: "Summarising en-net online forum statistics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Summarising en-net online forum statistics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(ennet)
library(dplyr)
```

The `ennet` package also includes analytic functions that summarises the text data available from the [**en-net** online forum](https://www.en-net.org). Currently, there are four sets of analytic functions available from `ennet`:

* `count_topics_time` functions - set of functions to count the number of topics or questions by day, wee, month or year;

* `count_topics_theme` functions - set of functions to count the number of topics or questions by theme;

* `count_topics_author` functions - set of functions to count the number of topics or questions by author; and,

* `arrange` functions - set of functions to arrange topics by number of views or by number of replies.

## Counting number of topics/questions

Summarising the number of topics or questions raised within the [**en-net** online forum](https://www.en-net.org) is basic and useful analytics that can proxy relative importance of a thematic area within the forum. This is facilitated by three sets of functions - 1) `count_topics_time`; 2) `count_topics_theme`; and, 3) `count_topics_author`.

### Counting number of topics/questions by time

The `count_topics_time` set of functions consist of four functions that count topics by *day*, by *week*, by *month*, or by *year*. To count the number of topics/questions per day, the `count_topics_day` function is used as follows:

```{r count_day1, echo = TRUE, eval = FALSE}
count_topics_day()
```

which results in:

```{r count_day1a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  count_topics_day()
```

By default, the `count_topics_day` function will provide an output that is not sorted by number of topics/questions per day. To sort by number of topics/questions per day, specify `.sort` as `TRUE`:

```{r count_day2, echo = TRUE, eval = FALSE}
count_topics_day(.sort = TRUE)
```

which gives the following output:

```{r count_day2, echo = FALSE, eval = TRUE}
ennet_topics %>%
  count_topics_day(.sort = TRUE)
```

Counting topics/questions by week, month, or year require just using the function named after the time required. For weekly counts, use `count_topics_week`, for monthly use `count_topics_month`, and for yearly use `count_topics_year`.

```{r count_time, echo = TRUE, eval = FALSE}
## Count topics/questions by week
count_topics_week()

## Count topics/questions by month
count_topics_month()

## Count topics/questions by year
count_topics_year()
```

### Counting topics/questions by theme

The `count_topics_theme` set of functions consists of two functions that count topics/questions *by theme*, and count topics/questions *by theme and by time*.

Counting of topics by theme is specified as follows:

```{r count_theme1, echo = TRUE, eval = FALSE}
count_topics_theme()
```

which results in:

```{r count_theme1a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  count_topics_theme()
```

Results provided by `count_topics_theme` is not sorted by number of topics/questions. To sort, specify `.sort` as `TRUE`:

```{r count_theme2, echo = TRUE, eval = FALSE}
count_topics_theme(.sort = TRUE)
```

which results in:

```{r count_theme2a, echo = FALSE, eval = TRUE}
count_topics_theme(.sort = TRUE)
```



Counting of topics overall across the years is specified as follows:

```{r usage8, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  count_topics(by_date = "all")
```

which results in:

```{r usage8a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  count_topics(by_date = "all")
```

By default, the output of `count_topics` arranges the results by descending frequency of counts of each topic by the grouping variable. This default behaviour can be changed by setting the `.sorted` argument to `FALSE`.

## Counting number of topics/questions raised by an author

Summarising the number of topics or questions raised within the [**en-net** online forum](https://www.en-net.org) for each author is another basic and useful analytics that can proxy level of engagement by those participating in the forum. This is facilitated using the `count_authors` function. Counting of topics for each author is done per thematic area and by a specific time period. Counting of topics for each author by month and year is performed by default:

```{r usage9, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  count_authors()
```

which results in:

```{r usage9a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  count_authors()
```

Counting of topics by year is specified as follows:

```{r usage10, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  count_authors(by_date = "year")
```

which results in:

```{r usage10a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  count_authors(by_date = "year")
```

Counting of topics overall across the years is specified as follows:

```{r usage11, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  count_authors(by_date = "all")
```

which results in:

```{r usage11a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  count_authors(by_date = "all")
```

By default, the output of `count_authors` arranges the results by descending frequency of counts of each topic raised by an author by the grouping variable. This default behaviour can be changed by setting the `.sorted` argument to `FALSE`.

## Arranging topics by number of views

Summarising the number of topics or questions raised within the [**en-net** online forum](https://www.en-net.org) by arranging them based on number of views can proxy level of interest to a specific topic by those participating in the forum. This is facilitated using the `arrange_views` function. Ranking of topics by number of views is done per thematic area and by a specific time period. Ranking of topics by number of views by thematic area and by month and year is performed by default:

```{r usage12, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_views()
```

which results in:

```{r usage12a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_views()
```

Arranging topics by number of views by thematic area and by year is performed as follows:

```{r usage13, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_views(by_date = "year")
```

which results in:

```{r usage13a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_views(by_date = "year")
```

Arranging topics by number of views by thematic area overall across the years is performed as follows:

```{r usage14, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_views(by_date = "all")
```

which results in:

```{r usage14a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_views(by_date = "all")
```

By default, the output of `arrange_views` is grouped by thematic area. This default behaviour can be changed by setting the `by_theme` argument to `FALSE`. For example, to arrange the topics by number of views by month and year across all themes:

```{r usage15, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_views(by_theme = FALSE)
```

which results in:

```{r usage15a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_views(by_theme = FALSE)
```

## Arranging topics by number of replies

Summarising the number of topics or questions raised within the [**en-net** online forum](https://www.en-net.org) by arranging them based on number of replies can proxy level of interest to a specific topic by those participating in the forum specifically those who provide responses and feedback to responses within the discussion. This is facilitated using the `arrange_replies` function. Ranking of topics by number of replies is done per thematic area and by a specific time period. Ranking of topics by number of replies by thematic area and by month and year is performed by default:

```{r usage16, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_replies()
```

which results in:

```{r usage17a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_replies()
```

Arranging topics by number of replies by thematic area and by year is performed as follows:

```{r usage18, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_replies(by_date = "year")
```

which results in:

```{r usage18a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_replies(by_date = "year")
```

Arranging topics by number of replies by thematic area overall across the years is performed as follows:

```{r usage19, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_replies(by_date = "all")
```

which results in:

```{r usage19a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_replies(by_date = "all")
```

By default, the output of `arrange_replies` is grouped by thematic area. This default behaviour can be changed by setting the `by_theme` argument to `FALSE`. For example, to arrange the topics by number of replies by month and year across all themes:

```{r usage20, echo = TRUE, eval = FALSE}
get_themes() %>%
  get_themes_topics() %>%
  arrange_replies(by_theme = FALSE)
```

which results in:

```{r usage20a, echo = FALSE, eval = TRUE}
ennet_topics %>%
  arrange_replies(by_theme = FALSE)
```
